### 2021-01-18 chemistry/doc.tex
""

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{mhchem}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Limiting reagents},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Limiting reagents}
\author{}
\date{}

\begin{document}
\maketitle

\hypertarget{limiting-reagents}{%
\section{Limiting reagents}\label{limiting-reagents}}

\hypertarget{lab-about-limiting-reagents}{%
\subsection{Lab about limiting
reagents}\label{lab-about-limiting-reagents}}

\(\ce{ZnO}\):\\
6.O1g before 4.10g after 1. Mass of ZnO used:
\[6.11 g - 4.10 g = 2.01 g\] 2. Molar mass of ZnO:
\[65.38 g/mol + 15.99 g/mol = 81.38 g/mol\] 3. How many moles of zinc
reacted with the HCl? \[2.01 g \div 81.38 g/mol = 0.0246 mol\] 4.
Balance the equation for this experiment
\[\ce{ZnO(aq) + 2 HCl(aq) -> ZnCl2(aq) + H2O(l)}\] 5. Find the number of
moles of \(\ce{HCl}\) in
\[\ce{HCl} = 1.01g/mol + 35.45g/mol = 36.46g/mol\]
\[\ce{HCl} = 20 \div 36.46g/mol = 0.549 mol\] 6. Find the number of
moles of \(\ce{HCl}\) in \[\ce{HCl} = 0.549 mol * 50 = 27.45 mol\]

\end{document}