### 2021-01-29 research project/sections/Background.tex
""

\section{Background}

\subsection{Focus on two methods}

In this paper, one by French mathematician François Viète,and another 
supposedly discovered by Madhava of Sangamagrama, and rediscovered 
by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
comparison, the process for the original discovery of these methods
will be explained. The two methods in question are one based on 
the infinite series definition of an inverse trigonometric function 
and one where the value is derived using geometry, by mathematician 
Viète. These two mathematical methods were chosen as they are both 
represent a first occurence in mathematics: Madhava was the first to 
find the series notation of the arctangent function and Viète was 
the first mathematician to use and infinite product in calculation. Furthermore, 
a research based on computational speed of two different kinds approaches to 
the constant $\pi$ has not been done to date.

\subsubsection{Madhava-Leibniz method}

Madhava is the first mathematician known to have found the series notation for 
the inverse tangent function, in the Kerala school of mathematics of Medieval India 
\cite{amermathmonthly}. One of

\subsubsection{Viète's method}

François Viète approached the value of $\pi$ from a 
geometric standpoint, and found an infinite product. 
He was able to calculate $\pi$ to a place of 9 decimal points, in 
the year 1593 \cite{Kreminski}, using his method. His method is reminiscent 
of Archimedes' method, where the length of a side is calculated \cite{archimedespi}, 
but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant 
radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon 
tends toward the area of a circle. The geometric origin of this formula 
can be found using simple right-angle trigonometry, by first finding 
the lengths $OH$ and subsequently $BD$ in \ref{fig:image1}.
% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
With the radius of the circle with center $R = OB$,

$OH = R \cos{\alpha}$

and

$BD = 2BH = 2 R \sin{\alpha}$

Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$, 
where $p$ is the perimeter of the polygon and $a$ is the apothem, in this case $BD \cdot n$ 
and $OH$ respectively, let $A_{n}$ equal the area of 
the polygon with $n$ sides such that: \\
$A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
$A_{n} = \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$

And if n is multiplied by 2, the angle $\angle \alpha$ is divided by 2, 
and the new area becomes:

$A_{2n} = 2n R^2 \sin{\frac{\alpha}{2}} \cos{\frac{\alpha}{2}}$

So it can be written that, by definition, the ratio of 
the area of an $n$-gon to one of a $2n$-gon is

$\frac{A_{n}}{A_{2n}} = \frac{n R^2 \sin{\alpha} \cos{\alpha}}{2 n R^2 \sin{\frac{\alpha}{2}} \cos{\frac{\alpha}{2}}} = \frac{\sin{2 \alpha}}{2 \sin{\alpha}}$

Which through the trigonometric identity $\sin{2 \theta} = 2\sin{\theta} \cos{\theta}$ can 
be simplified to:

$\frac{A_{n}}{A_{2n}} = \frac{2\sin{\alpha} \cos{\alpha} }{2\sin{\alpha}}  = \cos{\alpha}$

It can be then written that, through a new variable $P$, \\
$P = \frac{A_{n}}{A_{2n}} \frac{A_{2n}}{A_{4n}} \frac{A_{4n}}{A_{8n}} \dots \frac{A_{(k-2)n}}{A_{kn}} \frac{A_{kn}}{A}$ \\
where $A$ is the area of the circle of radius $R$ 
in \ref{fig:image1}.

So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and 
followingly, it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The 
value of $R = 1$ in this case, and since the area of a circle is defined 
by $A = \pi R^2$, therefore: \\
$\pi = \frac{A_{n}}{P}$ \\
By definition, we can say that as the value $k$ approaches infinity, the area of the $kn$-gon 
aproaches that of a circle, and therefore, the value of $\pi$.

$\frac{A_{n}}{\cos{\alpha} \cos{\frac{\alpha}{2}} \cos{\frac{\alpha}{4}} \dots}
\to_{k \to \infty} \pi$

Where the value of $A_{n}$ is the area of the first polygon, with $n=4$ sides, and as such 
$A_{n} = 4 \sin{45} \cos{45} = 2$. We can define: \\
$U_{0} = \cos{a} = \cos{45} = \frac{1}{\sqrt{2}}$ \\
$U_{1} = cos{\frac{\alpha}{2}}$ \\
Which we can, through the trigonometric identity 
$\cos^2{\theta} = \frac{1}{2} + \frac{1}{2} \cos{2\theta}$, simplify as
$U_{1} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{0}}$ \\
So it can be said that $U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$, which leads to 
a fully defined expression for the value of pi, using an infinite product:

$\pi = \frac{2}{\prod\limits_{k=0}^\infty U_{k}}, U_{0} = \frac{1}{\sqrt{2}}, U_{n} = \sqrt{\frac{1}{2} + \frac{1}{2} U_{n - 1}}$ \footnotemark

These expressions, when under a single expression result in the aforementioned 
formula with nested roots.

\footnotetext{where $\prod$ signifies a product. Similar expression to $\sum$ }

\begin{figure}[!ht]
\definecolor{qqwuqq}{rgb}{0,0.39215686274509803,0}
\definecolor{uuuuuu}{rgb}{0.26666666666666666,0.26666666666666666,0.26666666666666666}
\definecolor{xdxdff}{rgb}{0.49019607843137253,0.49019607843137253,1}
\definecolor{qqqqzz}{rgb}{0,0,0.6}
\begin{tikzpicture}[line cap=round,line join=round,>=triangle 45,x=1cm,y=1cm,scale=0.7]
\clip(-8.372520930232552,-8.50470697674418) rectangle (7.69259534883721,7.028316279069767);
\draw [shift={(-0.03355999206952884,0)},line width=1pt,color=qqwuqq,fill=qqwuqq,fill opacity=0.10000000149011612] (0,0) -- (0:1.3302325581395344) arc (0:44.77428433325932:1.3302325581395344) -- cycle;
\draw [line width=0.8pt,color=qqqqzz] (0,0) circle (6cm);
\draw [line width=1pt] (4.242640687119286,4.242640687119285)-- (6,0);
\draw [line width=1pt] (6,0)-- (4.242640687119286,-4.242640687119285);
\draw [line width=1pt] (4.242640687119286,-4.242640687119285)-- (4.242640687119286,4.242640687119285);
\draw [line width=1pt] (-6,0)-- (6,0);
\draw [line width=1pt] (-4.275936263984878,-4.209081736713964)-- (4.242640687119286,4.242640687119285);
\draw [line width=1pt] (-4.242640687119285,4.242640687119286)-- (4.242640687119286,-4.242640687119285);
\draw [line width=1pt] (-0.03355999206952884,0)-- (5.121320343559643,2.1213203435596424);
\draw [line width=1pt] (-4.242640687119285,4.242640687119286)-- (-4.275936263984878,-4.209081736713964);
\begin{scriptsize}
\draw [fill=xdxdff] (4.242640687119286,4.242640687119285) circle (2.5pt);
\draw[color=xdxdff] (4.4772465116279085,4.623665116279071) node {$B$};
\draw [fill=xdxdff] (-4.242640687119285,4.242640687119286) circle (2.5pt);
\draw[color=xdxdff] (-4.295311627906972,4.623665116279071) node {$A$};
\draw [fill=xdxdff] (-4.275936263984878,-4.209081736713964) circle (2.5pt);
\draw[color=xdxdff] (-4.336241860465112,-4.55) node {$C$};
\draw [fill=xdxdff] (6,0) circle (2.5pt);
\draw[color=xdxdff] (6.237246511627907,0.3873860465116302) node {$E$};
\draw [fill=xdxdff] (4.242640687119286,-4.242640687119285) circle (2.5pt);
\draw[color=xdxdff] (4.3772465116279085,-4.5) node {$D$};
\draw[color=black] (5.3595720930232575,2.3111069767441874) node {$H'$};
\draw [fill=xdxdff] (5.12,2.12) circle (2.5pt);
\draw[color=black] (3.947479069767444,-0.615404651162788) node {$H$};
\draw [fill=xdxdff] (4.242640687119286,0) circle (2.5pt);
\draw [fill=xdxdff] (0,0) circle (2.5pt);
\draw[color=uuuuuu] (0.00003720930232879,-0.3464558139534907) node {$O$};
\draw[color=qqwuqq] (1.5325953488372117,0.3873860465116302) node {$\alpha$};
\end{scriptsize}
\end{tikzpicture}
\label{fig:image1}
\caption{Circle with 1 segment from a $n$-gon with point $H$ and 2 
segments from an $2n$-gon, one of which on point $H'$, inscribed in 
a circle of radius $OB$, adapted from (Boris Gourévitch) \cite{borisgourevitch2013}}
\end{figure}