### 2021-01-29 research project/sections/Computational_approach.tex
workedo n rp. added graphs and tables

\section{Computational approach}

\subsection{The variables}
The dependent variable of this experiment is the time taken $t$ by the program 
to approximate a given number $n$ of correct decimal value of the constant $\pi$.

The value $n$ will be altered in order to avoid possible similar convergence rates 
at a small amount of decimal places, and multiple trials will be run to decrease 
margin of error. Other variables of the experiment will be controlled. For example, the experiment 
will be run on a same isolated system, a virtual machine, with a minimal amount of 
processes running to avoid any possible variance in results.

\subsection{Implementing in Python}

A Python application was programmed (see appendix) in order to run the two 
methods aforementioned, and manage the collection of data.

The program made assigns the time before the execution of the method to a variable \verb|t1| 
with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal 
places of the resultant approximation are counted and once a specified threshold is reached, 
a new \verb|t2| time variable is assigned and the time taken, defined by the difference between 
\verb|t2| and \verb|t1|, is stored. This process is repeated for all specified decimal accuracies and 
for both methods. The times recorded were stored in a \verb|.csv| file for further analysis.

The \verb|mpmath| library was used for the floating point operations required 
for comparison between approximated values and the constant $\pi$ that wouldn't 
have been possible using standard Python libraries \cite{mpmath}.