### 2021-01-31 research project/sections/Further_research.tex
rp

\section{Further research opportunities}

\subsection{}