### 2021-02-01 research project/sections/Further_research.tex

\section{Further research opportunities}

\subsection{Investigating newer, more efficient algorithms}

As the computation of $\pi$ moved from paper to machines, as all did from accounting to 
mathematics, newer and more efficient methods were developed, notably by Ramanujan and 
the Chudnovsky brothers.