### 2021-02-02 research project/sections/Analysis_of_results.tex
rp

\section{Analysis of the results}
\newcolumntype{C}{>{\Centering\arraybackslash\hspace{0pt}}X}

\pgfplotstableread{
    X Y
    5 0.7601666451
    10 1.288158894
    15 2.050116062 
    20 2.63982296 
    25 3.091802597
    30 3.819538355 
    35 4.442822933
    40 5.081979036 
    45 5.706284046 
    50 6.370962858 
    55 7.020516396 
    60 7.723451853   
}\vietetable

\pgfplotstableread{
    X Y
    5 0.5919837952 
    10 1.307063103  
    15 1.902475357  
    20 2.679748535  
    25 3.212277889  
    30 3.904948235  
    35 4.547913074  
    40 5.255579948  
    45 5.837199688  
    50 6.578524113  
    55 7.270073891  
    60 8.004565239

}\madhavatable

\pgfplotstableread{
    X Y
    5 10
    10 17
    15 27
    20 35
    25 42
    30 51
    35 59
    40 67
    45 76
    50 84
    55 92
    60 101
}\vietetableiter

\pgfplotstableread{
    X Y
    5 9
    10 20
    15 29
    20 41
    25 50
    30 61
    35 71
    40 82
    45 91
    50 102
    55 112
    60 123
}\madhavatableiter

\subsection{Presentation of the data}

\subsubsection{Tabular presentation}

The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
in the table shown below were found (rounded to 3 significant figures). The values for the decimal place 
0 have been dropped as it would not be logical to include them.

\begin{table}[h]
    \noindent% <-- important
    \setlength\tabcolsep{3pt} % default: 6pt
    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
    \toprule
    Decimal place & Average time for approximation using Viete's method, in milliseconds (ms) & Average time for approximation using Madhava's method method, in milliseconds (ms) \\ 
    \midrule
    5  & 0.760  & 0.592   \\
    10 & 1.29  & 1.31    \\
    15 & 2.05  & 1.9     \\
    20 & 2.64  & 2.68    \\
    25 & 3.09  & 3.21    \\
    30 & 3.82  & 3.9     \\
    35 & 4.44  & 4.55    \\
    40 & 5.08  & 5.26    \\
    45 & 5.71  & 5.84    \\
    50 & 6.37  & 6.58    \\
    55 & 7.02  & 7.27    \\
    60 & 7.72  & 8.00   
    \end{tabularx}
\end{table}

\begin{table}[h]
    \noindent% <-- important
    \setlength\tabcolsep{3pt} % default: 6pt
    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
    \toprule
    Decimal place & Iterations needed, Viète's method (no unit) & Iterations needed, Madhava's method (no unit)\\
    \midrule
    5 & 10 & 9 \\
    10 & 17 & 20 \\
    15 & 27 & 29 \\
    20 & 35 & 41 \\
    25 & 42 & 50 \\
    30 & 51 & 61 \\
    35 & 59 & 71 \\
    40 & 67 & 82 \\
    45 & 76 & 91 \\
    50 & 84 & 102 \\
    55 & 92 & 112 \\
    60 & 101 & 123
    \end{tabularx}
\end{table}

The iterations needed for both methods were also measured, using the Python program 
(see appendix), slightly modified to increment a variable \verb|i| and subsequently 
return it, writing it in a similar manner to a \verb|.csv| file.

\subsubsection{Graphical presentation}

Shown below is a graph demonstrating the correlation between the decimal places approximated 
and the amount of iterations needed for this specific decimal milestone. The blue markers and 
line of best-fit respresent the results received from Viète's method while the red represent those 
from Madhava's method. \\

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
        title={Comparing the iterations needed for approximating the value of $\pi$},
        xlabel={Decimal places},
        ylabel={Iterations},
        xmin=0, xmax=60,
        ymin=0, ymax=130,
        ytick={0,10,20,30,40,50,60,70,80,90,100,110,120,130},
        xtick={0,10,20,30,40,50,60},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        scale=1.2
    ]

    \addplot[
        only marks,
        color=blue,
        mark=*,
        ] table {\vietetableiter};
        \legend{Viète's method}

    \addplot [thin, blue] table[
        y={create col/linear regression={y=Y}}
    ] % compute a linear regression from the input table
    {\vietetableiter};
    \addlegendentry{best fit line}

    \addplot[
        only marks,
        color=red,
        mark=*,
        ] table {\madhavatableiter};
        \addlegendentry{Madhava's method}

    \addplot [thin, red] table[
        y={create col/linear regression={y=Y}}
    ] % compute a linear regression from the input table
    {\madhavatableiter};
    \addlegendentry{best fit line}

    \end{axis}
\end{tikzpicture}

To show the trend in the time data collected, see below the decimal places compared 
to the time taken per method. The color coding here is relevant to the one aforementioned. \\

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
        title={Comparing the time taken for approximating the value of $\pi$},
        xlabel={Decimal places},
        ylabel={Time taken [ms]},
        xmin=0, xmax=60,
        ymin=0, ymax=10,
        ytick={0,2,4,6,8,10},
        xtick={0,10,20,30,40,50,60},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        scale=1.2
    ]

    \addplot[
        only marks,
        color=blue,
        mark=*,
        ] table {\vietetable};
        \legend{Viète's method}

    \addplot [thin, blue] table[
        y={create col/linear regression={y=Y}}
    ] % compute a linear regression from the input table
    {\vietetable};
    \addlegendentry{best fit line}

    \addplot[
        only marks,
        color=red,
        mark=*,
        ] table {\madhavatable};
        \addlegendentry{Madhava's method}

    \addplot [thin, red] table[
        y={create col/linear regression={y=Y}}
    ] % compute a linear regression from the input table
    {\madhavatable};
    \addlegendentry{best fit line}

    \end{axis}
\end{tikzpicture}

\subsection{Observations and analysis}

The results gathered from this experiment show linear relationship between the amount of decimal 
places approximated and the time required for this approximation, as well as between the decimal places 
and the amount of iterations needed. The two lines of best fit for the two methods show 
similarities but also differences.

Where decimal places and iterations are compared, there is a clear difference between the two 
methods: Madhava's method indeed takes more iterations to converge to a specific decimal of 
the value $\pi$. However, in the second graph comparing time, it can be noted that Madhava's 
method is marginally faster until the decimal 20, when the two methods diverge. It could be 
assumed that this occurrence is due to differences between the types of the two methods. Madhava's 
method is what is called an alternating series, due to its nature of alternating between values in order 
to converge to a specific value, in this case $\pi$, while Viète's method is an infinite product. The 
difference between these two methods is underlined in (Figure 3) \ref{fig:comparaison}. It can be 
deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation 
accuracy increases in what looks like an exponential fashion.

\textit{More info to be added}

\begin{figure}[h]
    \includegraphics[width=\linewidth]{image.png}
    \caption{Comparison of historical methods of approximating the value of $\pi$. This image demonstrates
    their convergence rates. The line labelled Madhava in dark blue is the method used in this paper. 
    From (Wikimedia Commons) \cite{infinite_series_comparaison}}
    \label{fig:comparaison}
  \end{figure}