### 2021-02-19 math/ia_2021/paper/paper.tex
math ia

% !TEX root = paper.tex

\documentclass[12pt,a4paper]{article}

% Packages
\usepackage{packages}

% Document Settings
\geometry{margin=1in}
\setlength{\parskip}{1ex}
\setlength{\parindent}{0pt}

% =========================================
%             DOCUMENT
% =========================================

\begin{document}
\doublespacing % Double spacing throughout the document

% =========================
%      TITLE PAGE
% =========================

% Suppresses headers, footers, and page numbers on title page
\begin{titlepage}
    \begin{center}
    \vspace*{4cm}
    A statistical approach to predicting a tsunami’s wave height \\
    \vspace{1cm}
    Can the height of a tsunami's wave be predicted based on the distance from and 
    the magnitude of a preceding earthquake? \\
    \vspace{4cm}
    Word count: TBD
    \vfill
    \vspace{0.1cm}
    \end{center}
    \end{titlepage}

% =========================
%      DOCUMENT BODY
% =========================

\pagenumbering{roman}

\begin{center}
\pdfbookmark{\contentsname}{Contents}
\tableofcontents
\vspace{1in}

\end{center}

% TODO: Add hyper-references to links
% TODO: Improve listings' styling
% TODO: Add links to Contents entries

\newpage

\pagenumbering{arabic}

\include{sections/Introduction}

\end{document}