### 2021-03-07 research project/sections/Computational_approach.tex

\section{Computational approach}

\subsection{The variables}
The dependent variable of this experiment is the time taken $t$ by the program
to approximate a given number $n$ of correct decimal value of the constant $\pi$.

The value $n$ will be altered in order to avoid possible similar convergence rates
at a small amount of decimal places, and multiple trials will be run to decrease
margin of error. Other variables of the experiment will be controlled. For example, the experiment
will be run on a same isolated system, a virtual machine, with a minimal amount of
processes running simultaneously to avoid any possible variance in results.

\subsection{Implementing in Python}

A Python application was programmed (see appendix) in order to run the two
methods aforementioned, and manage the collection of data.

The program assigns the time before the execution of the method to a variable \verb|t1|
with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal
places of the resultant approximation are counted. Once a specified decimal place is reached,
determined by a programming \verb|if| statement, that can be mathematically expressed as:

$| \pi - \pi_{approx} | < 10^{-n}$, where $\pi_{approx}$ is the approximation by the method and 
$n$ defines the number of correct decimal places desired.

A new \verb|t2| time variable is assigned and the time taken, defined by $\verb|t2| - \verb|t1|$, is stored. This process is repeated for all specified decimal accuracies and
for both methods. The times recorded were stored in a \verb|.csv| file for further analysis.

The \verb|mpmath| library was used for the floating point operations required
for comparison between approximated values and the constant $\pi$ that wouldn't
have been possible using standard Python libraries, due to floating-point limitations (computing 
term for decimal numbers). \cite{mpmath}