### 2021-03-07 research project/sections/Conclusion.tex

\section{Conclusion}

To conclude, even though Viète's method can seem more convoluted in practice, due to its geometric origins,
it can be seen through the experiment conducted in this paper that this is clearly not the case
in a computational context. His method is simplified to a point where its geometric origins are no
longer findable from its computational implementation. And despite the assumption that could be
made based on the types of operations needed for each method, Viète's requiring arguably more
computationally challenging operations. These operations, such as square-root function or the product function
used on high-precision floating point numbers, can be said to be more demanding than what could be said
is simple summation, in the case of Madhava.

An example of this kind of difference in computing power 
needed for different mathematical operations can be seen in the video-game \textit{Quake III Arena}, published in 1999. 
The inverse square root of a number needed to be calculated, however due to performance issues, the standard 
function could not be used. And as such, a new function for the approximation of the inverse square root of a number 
was programmed \cite{id-software_2012}, highly more efficient than the old one, which shows that a mathematical function 
implemented in computing can be more efficient than another.

Ultimately, a method 
of approximating the constant $\pi$ computationally can be more efficient than another, and to the 
somewhat surprising extent that this method is a geometric one --- many would think an analytical one, 
that uses simple summation as opposed to square roots, would be more efficient.

\subsection{Evaluation of the experimental method}

The method used in this paper is not perfect, and at closer inspection and in 
retrospect, many issues can be found. Firstly, in the computational implementation of 
Madhava's method, the value $\sqrt{12}$ is calculated at every iteration, in order to 
determine the number of correct decimal places, which can 
lead to efficiency losses. Instead of this, the value could have been calculated once 
and stored in a variable, to be referenced at every iteration. However, after running 
the program again with the modified code, the results stayed similar.

Another weakness in the method used in this paper is that a higher number of trials 
could have been used, as well as the value of $\pi$ approximated to a higher 
number of decimal places. A higher number of decimal places was disregarded during the 
implementation of the methods as it was noticed that values over 60 would take an indeterminately 
long time. This should not have been overlooked, as a different kind of relationship 
could have been made apparent that way.