### 2021-03-07 research project/sections/Further_research.tex

\section{Further research opportunities}

\subsection{Investigating newer, more efficient algorithms}

As the computation of $\pi$ moved from paper to machines, as all did from accounting to
mathematics, newer and more efficient methods were developed, notably by Ramanujan and
the Chudnovsky brothers. These methods should also be investigated, as they 
are arguably more pertinent in this field.

\subsection{Investigating the uses of efficient code}

This paper does not discuss the many advantages of efficient code in detail, 
and a research on the speed of the mathematical functions in programming could 
be conducted. Computer program efficiency is always a pertinent topic because 
of a faster and faster world. An example of this could be the company Amazon, 
whose production speed has skyrocketed during the last decade. An increasingly 
fast demand also requires an increasingly fast computing infrastructure; to handle 
shipments, manage customers or even conduct job interviews (an experimental system that received 
major backlash in 2018). \cite{murad_2021}